---
output: html_document
---
Reproducible Research Course Project 1
==========================


Load the activity.csv file
```{r}
library(dplyr)
library(knitr)
data = read.csv("activity.csv")

```

1) What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day:
```{r}

op = par(mfrow=c(1,1), mar=c(4,4,2,2))
with(data, hist(steps, main="Total Steps per Day",col= rainbow(5))); op

```

Calculate and report the mean and median total number of steps taken per day
```{r}

summarise(data, mean = mean(steps, na.rm = TRUE))
summarise(data, median = median(steps, na.rm = TRUE))

```

2) What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}

groupby <- group_by(data, date)
summarise(groupby, mean = mean(steps))

```

```{r}

op = par(mfrow=c(1,1), mar=c(4,4,2,2))
with(avgAll,{
        plot(date, mean, type="n", main="Average Steps", xlab="Day of the Week", ylab="Mean")
        lines(date,mean)
})
op

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}

groupBy3 <- group_by(data, date)
avgAll2 = summarise(groupBy3, mean = mean(steps))
max = max(avgAll2$mean, na.rm = TRUE); max

```

3) Imputing missing values

Devise a strategy for filling in all of the missing values in the dataset. 
```{r}

avgAll3[is.na(avgAll3)] = 0

```


Make a histogram of the total number of steps taken each day  

```{r}
op = par(mfrow=c(2,1), mar=c(4,4,2,2))
hist(avgAll2$mean,
     main="With Missing Values",
     xlab = "Average Steps",
     col= rainbow(5))
hist(avgAll3$mean,
     main="Missing Values Replaced with 0",
     xlab = "Average Steps",
     col= rainbow(5))      
op


```


4) Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
addWeekdays =
mutate(avgAll3,
      weekdays = ifelse(!weekdays(as.Date(avgAll3$date)) %in% c("Saturday", "Sunday"),"weekday","weekend"))

filterWeekdays = filter(addWeekdays, weekdays == 'weekday')       
filterWeekends = filter(addWeekdays, weekdays == 'weekend')

weekdayDate = filterWeekdays$date
weekendDate = filterWeekends$date

weekdayMean = filterWeekdays$mean
weekendMean = filterWeekends$mean

```

Make a panel plot containing a time series plot

```{r}
op = par(mfrow=c(2,1), mar=c(4,4,2,2))
with(filterWeekdays, {        
        plot(weekdayDate, weekdayMean, type="n", main="Weekdays", ylab="Mean", xlab="Day of the Week") 
        lines(weekdayDate, weekdayMean)
})
with(filterWeekends, {        
        plot(weekendDate, weekendMean, type="n", main="Weekends", ylab="Mean", xlab="Day of the Week") 
        lines(weekendDate, weekendMean)
}) 
op

```



