---
title: "BMI calculator"
author: "Michael Malin Jr"
date: "Saturday, August 08, 2015"
output: html_document
runtime: shiny
---

```{r echo = FALSE}
bmiApp <- function(feet, inches, weight) {
        require(shiny)
        shinyApp(
                ui = pageWithSidebar(
                        # Application title
                        headerPanel("Body Mass Index(BMI) calculator"),
                        sidebarPanel(
                                numericInput('feet', 'Feet', 5, min = 3, max = 7, step = 5),
                                numericInput('inches', 'Inches', 5, min = 1, max = 12, step = 5),
                                numericInput('weight', 'Weight', 150, min = 75, max = 300, step = 5),
                                submitButton('Submit')
                        ),
                        mainPanel(
                                h3('Results:'),                                
                                h4('Your BMI is:'),
                                verbatimTextOutput("inputValue"),                                
                                h4('Weight Category:'),
                                verbatimTextOutput("prediction")
                                )
                        ),                 
               
                server = function(input, output) {
                        # BMI calculation
                        bmi = function(feet, inches, weight){
                                w = weight * 703
                                h.Inches = (feet * 12) + inches
                                h.square = h.Inches * h.Inches
                                calc = w/h.square  
                                print(calc)
                        }
                        bmi.pred = function(feet, inches, weight){
                                        w = weight * 703
                                        h.Inches = (feet * 12) + inches
                                        h.square = h.Inches * h.Inches
                                        calc = w/h.square
                                        
                                        if(calc <= 18.5){
                                                print("Underweight")                        
                                        } else if(calc >= 18.5 & calc <= 24.9){
                                                print("Normal weight")                        
                                        } else if(calc >= 25 & calc <= 29.9){
                                                print("Overweight")                        
                                        } else{
                                                print("Obesity")}
                                        
                                }
                        output$inputValue <- renderPrint({bmi(input$feet, input$inches, input$weight)})
                        output$prediction <- renderPrint({bmi.pred(input$feet, input$inches, input$weight)})
                }
                
        )        
        
}
```

## Body Mass Index(BMI) calculator

This paper will show you how to calculate your BMI and your weight category with my bmiApp() function. 
BMI is used to help you measure if you are underweight, normal weight, overweight, or obese
based on your height and weight. 

###### You can run the app with the following function:

```{r echo = TRUE}
bmiApp()
```

###### Below is the code for the user interface(ui), which controls how the app looks, and the server, which controls what the app does: 

```{r echo = TRUE}
bmiApp <- function(feet, inches, weight) {
        require(shiny)
        shinyApp(
                ui = pageWithSidebar(
                        # Application title
                        headerPanel("Body Mass Index(BMI) calculator"),
                        sidebarPanel(
                                numericInput('feet', 'Feet', 5, min = 3, max = 7, step = 5),
                                numericInput('inches', 'Inches', 5, min = 1, max = 12, step = 5),
                                numericInput('weight', 'Weight', 150, min = 75, max = 300, step = 5),
                                submitButton('Submit')
                        ),
                        mainPanel(
                                h3('Results:'),                                
                                h4('Your BMI is:'),
                                verbatimTextOutput("inputValue"),                                
                                h4('Weight Category:'),
                                verbatimTextOutput("prediction")
                                )
                        ),                 
               
                server = function(input, output) {
                        # BMI calculation
                        bmi = function(feet, inches, weight){
                                w = weight * 703
                                h.Inches = (feet * 12) + inches
                                h.square = h.Inches * h.Inches
                                calc = w/h.square  
                                print(calc)
                        }
                        bmi.pred = function(feet, inches, weight){
                                        w = weight * 703
                                        h.Inches = (feet * 12) + inches
                                        h.square = h.Inches * h.Inches
                                        calc = w/h.square
                                        
                                        if(calc <= 18.5){
                                                print("Underweight")                        
                                        } else if(calc >= 18.5 & calc <= 24.9){
                                                print("Normal weight")                        
                                        } else if(calc >= 25 & calc <= 29.9){
                                                print("Overweight")                        
                                        } else{
                                                print("Obesity")}
                                        
                                }
                        output$inputValue <- renderPrint({bmi(input$feet, input$inches, input$weight)})
                        output$prediction <- renderPrint({bmi.pred(input$feet, input$inches, input$weight)})
                }
                
        )        
        
}
```


#### References:

1) http://www.whathealth.com/bmi/formula.html
2) http://www.nhlbi.nih.gov/health/educational/lose_wt/BMI/bmicalc.htm

