---
title: "Body Mass Index (BMI) Calculator"
author: "Michael Malin Jr"
date: "Monday, August 10, 2015"
output: html_document
---

## Calculating Body Mass Index

 * The Body Mass Index formula was created by astronomer Adolphe Quetelet.
 * As obesity increased in Western societies, BMI used as the best measure of body fat.


## Introducing bmiApp()

* BMI is used to help you measure if you are underweight, normal weight, overweight, or obese based on your height and weight. 
* I will show you how to calculate your BMI and your weight category with my bmiApp() function.
 
        
## How The App Works
        
* The following [link](https://github.com/michaelmalinjr/datasciencecoursera/blob/master/Data%20Products/bmiApp.R) to the code for the user interface(ui), which controls how the app looks, and the server, which controls what the app does. 

        
## App In Motion

```{r echo = FALSE}
bmiApp <- function(feet, inches, weight) {
        require(shiny)
        shinyApp(
                ui = pageWithSidebar(
                        # Application title
                        headerPanel("Body Mass Index(BMI) calculator"),
                        sidebarPanel(
                                numericInput('feet', 'Feet', 5, min = 3, max = 7, step = 5),
                                numericInput('inches', 'Inches', 5, min = 1, max = 12, step = 5),
                                numericInput('weight', 'Weight', 150, min = 75, max = 300, step = 5),
                                submitButton('Submit')
                        ),
                        mainPanel(
                                h3('Results:'),                                
                                h4('Your BMI is:'),
                                verbatimTextOutput("inputValue"),                                
                                h4('Weight Category:'),
                                verbatimTextOutput("prediction")
                        )
                ),                 
                
                server = function(input, output) {
                        # BMI calculation
                        bmi = function(feet, inches, weight){
                                w = weight * 703
                                h.Inches = (feet * 12) + inches
                                h.square = h.Inches * h.Inches
                                calc = w/h.square  
                                print(calc)
                        }
                        bmi.pred = function(feet, inches, weight){
                                w = weight * 703
                                h.Inches = (feet * 12) + inches
                                h.square = h.Inches * h.Inches
                                calc = w/h.square
                                
                                if(calc <= 18.5){
                                        print("Underweight")                        
                                } else if(calc >= 18.5 & calc <= 24.9){
                                        print("Normal weight")                        
                                } else if(calc >= 25 & calc <= 29.9){
                                        print("Overweight")                        
                                } else{
                                        print("Obesity")}
                                
                        }
                        output$inputValue <- renderPrint({bmi(input$feet, input$inches, input$weight)})
                        output$prediction <- renderPrint({bmi.pred(input$feet, input$inches, input$weight)})
                }
                
        )        
        
}
```

```{r echo = FALSE}
bmiApp()
```
        
        
## References
        
1. [www.whathealth.com](http://www.whathealth.com/bmi/formula.html)

2. [www.nhlbi.nih.gov](http://www.nhlbi.nih.gov/health/educational/lose_wt/BMI/bmicalc.htm)

3. [Adolphe Quetelet Wiki](https://en.wikipedia.org/wiki/Adolphe_Quetelet)

4. [Body Mass Index Wiki](https://en.wikipedia.org/wiki/Body_mass_index)
